---
title: "Capstone Final Report"
author: "Samson Lei"
date: "May 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Capstone Introduction

"PLAYERUNKNOWN'S BATTLEGROUNDS (PUBG) is a competitive survival shooter. Players are dropped into a wide, open area, and they must fight to the death - all while the battlefield shrinks, adding pressure to all in its grip." -Reddit's about PUBG statement

For this capstone I will analyze player statistics from the video game "PlayerUnknown's Battlegrounds" aka PUBG. The original goal of this analysis was to recommend other players of this game possible strategies based on the datasets provided from posted on Kaggle, which originated from the PUBG developer's own data collecting methods. A side goal that would have contributed to that idea would have been to present outliers in the player statistics that denoted significant difference in skill or ability in-game. However, through a more in-depth look at the data we have, I believe that the same anaylsis could serve the developers of the game much better. 

As such, the primary goal of this analysis is to seek out notable outliers in the data. The objective of this goal is to provide insight into these players.

### Significance

The importance of the analysis of outliers can be placed into two major categories. 

1. Outliers displaying overall greater skill

* The developers/publishers for PUBG has been known to seek out players / content makers as partners for the game as a form of advertisement and to promote the game.
* Identification of players with skill who are also csontent makers is an important form of markerting for the game.
* Skilled players that may not be content makers might also hold the possibility of being significant in terms of tournament participation for the game, which furthers promotion for the game.

2. Outliers displaying strange behaviors and patterns

* Identifying cheating.
* Cleaning out player statistics that are primarily caused by custom matches which do not contribute to player data with any real value. 
* Custom match statistics tend to be overwhelmingly skewed due to the entertainment purpose of the match rather than actual gameplay.

## Overview of data sets/ files

The original dataset included a single file from Kaggle: 
* "PUBG_PLAYER_Statistics.csv"

Segregated files:
* "PUBG_Player_Statistics_solo"
* "PUBG_Player_Statistics_duo"
* "PUBG_Player_Statistics_squad"

I broke down the original dataset into 3 primary categories wherein each represent a team composition:

* Solo
* Duo (2 player team)
* Squad (3+ player team)

```{r, echo=FALSE}
load(file="PUBG_Player_Statistics_solo.RData")
summary(PUBG_Player_Statistics_solo)
```

In comparing the Maximum values to the mean values, we can already speculate that there are significant outliers in the datasets due to the difference in distance between the values. 

```{r, echo=FALSE}
load(file="PUBG_Player_Statistics_duo.RData")
summary(PUBG_Player_Statistics_duo)
```

```{r, echo=FALSE}
load(file="PUBG_Player_Statistics_squad.RData")
summary(PUBG_Player_Statistics_squad)
```

Similarly the same trend from the solo statistics are noticed within both duo and squad datasets.

## Data Wrangling

Within the datasets, cleaning was primarily focused on the solo statistics due to some of the column names not being utilized while playing alone.

Categories removed primarily from solo dataset:

* Revivals were removed since in solo play there are no teammates for you to revive.

* DBNO, which stands for "Down But Not Out", is not used in solo play either. In PUBG, a player only goes DBNO when they are on a team which can still save them before they completely die off. In PUBG, no matter what team composition, a player will die instantly if there are no surviving or not DBNO teammates.

Categories removed from all segregated datasets:

* TrackerID was also removed, since it wasn't a value to be analyzed.

* WeaponAcquired wasn't a very relevant statistic since every value was zero. 


### Secondary datasets

Luckily a vast majority of the data, aside from the player names, were numeric. No conversions were made to statistics in the datasets due to this. 

However I did make a significant breakdown of the data that involved the number of rounds played by each player. I had skimmed out the players who have played less than **200** rounds for this analysis. I believe that the players with another range of number of played rounds could also be analyzed further for a similar purpose as the current one, but I think that would take much too long to search through. Unfortunately this dataset is simply too vast to not cut down like this. 

Similar to the original idea of breaking down the datasets based upon team compositions:

### Solo dataset

```{r, echo=FALSE}
load(file="solo.RData")
summary(solo)
```

### Duo dataset

```{r, echo=FALSE}
load(file="duo.RData")
summary(duo)
```

### Squad dataset

```{r, echo=FALSE}
load(file="team.RData")
summary(team)
```



### Exploratory Data Analysis

**Although I do not think I will utilize some of the columns in this analysis I believe that information is still relevant enough that it should remain.**

```{r, echo=FALSE}
load(file="solo.RData")
colnames(solo)
```

```{r, echo=FALSE}
load(file="duo.RData")
colnames(duo)
```

```{r, echo=FALSE}
load(file="team.RData")
colnames(team)
```


Since we are seeking outliers, seeking out the maximum values of each column would be a good start.
However, since there are so many columns I will focus primarily on some main categories:

* **KillDeathRatio**: Common measure of skill (even if the player does not win the game overall).
* **WinRatio**: Win to loss ratio (May help identify cheaters as well as skilled players).
* **RoundsPlayed**: The number of rounds played is important to determine actual experience among players.
* **Wins**: Solid win count (since players can absolutely start off not very good at the game, their overall win count may help us see their growth overtime more than their win ratio).
* **Kills**: Solid kill count from all rounds (following a similar reasoning to K/D, it can also provide further insight on how well a player does in game.)

The remaining statistics can be useful in supporting the analysis, but would not be overall primary focuses. They will be used to help augment any notable trends. 

#### Notable players by solo statistics 
**Top players by kill death ratio**

```{r, echo=FALSE}
load(file="solo.RData")
head(solo[order(-solo$solo_KillDeathRatio),])
```

**Top players by win ratio**

```{r, echo=FALSE}
load(file="solo.RData")
head(solo[order(-solo$solo_WinRatio),])
```

**Players with the greatest amount of experience in terms of rounds played**

```{r, echo=FALSE}
load(file="solo.RData")
head(solo[order(-solo$solo_RoundsPlayed),])
```

**Players with the most wins**

```{r, echo=FALSE}
load(file="solo.RData")
head(solo[order(-solo$solo_Wins),])
```

**Players with the most kills**

```{r, echo=FALSE}
load(file="solo.RData")
head(solo[order(-solo$solo_Kills),])
```

#### Notable players by duo statistics 
**Top players by kill death ratio**

```{r, echo=FALSE}
load(file="duo.RData")
head(duo[order(-duo$duo_KillDeathRatio),])
```

**Top players by win ratio**
  
```{r, echo=FALSE}
load(file="duo.RData")
head(duo[order(-duo$duo_WinRatio),])
```

**Players with the greatest amount of experience in terms of rounds played**
  
```{r, echo=FALSE}
load(file="duo.RData")
head(duo[order(-duo$duo_RoundsPlayed),])
```

**Players with the most wins**
  
```{r, echo=FALSE}
load(file="duo.RData")
head(duo[order(-duo$duo_Wins),])
```

**Players with the most kills**
  
```{r, echo=FALSE}
load(file="duo.RData")
head(duo[order(-duo$duo_Kills),])
```


#### Notable players by team statistics 
**Top players by kill death ratio**

```{r, echo=FALSE}
load(file="team.RData")
head(team[order(-team$squad_KillDeathRatio),])
```

**Top players by win ratio**
  
```{r, echo=FALSE}
load(file="team.RData")
head(team[order(-team$squad_WinRatio),])
```

**Players with the greatest amount of experience in terms of rounds played**
  
```{r, echo=FALSE}
load(file="team.RData")
head(team[order(-team$squad_RoundsPlayed),])
```

**Players with the most wins**
  
```{r, echo=FALSE}
load(file="team.RData")
head(team[order(-team$squad_Wins),])
```

**Players with the most kills**
  
```{r, echo=FALSE}
load(file="team.RData")
head(team[order(-team$squad_Kills),])
```

As a preliminary search across the separated datasets, these players would stand out the most due to their statistics in each field. 

## Data exploration via visualizations

To further investigate outliers in the player statistics, we can also visualize the K/D ratios in a histogram to verify at what point a K/D becomes above average and notable. 

### K/D visualization

```{r, echo=FALSE}
load(file="solo.RData")

#Solo K/D histogram
qplot(solo$solo_KillDeathRatio, 
      xlab="Solo K/D ratio",
      geom="histogram",
      binwidth=.01, 
      fill=I("blue"), 
      col=I("red"))

#Duo K/D histogram
qplot(duo$duo_KillDeathRatio, 
      xlab="Duo K/D ratio",
      geom="histogram",
      binwidth=.01, 
      fill=I("blue"), 
      col=I("red"))

#Squad K/D histogram
qplot(team$squad_KillDeathRatio, 
      xlab="Squad K/D ratio",
      geom="histogram",
      binwidth=.01, 
      fill=I("blue"), 
      col=I("red"))

```

Utilizing the segregated dataset which has been further cut down to players with at least 200 rounds of gameplay experience, we can see that in this set a significant portion of experienced players have a K/D less than a 2.0. 

It is important to note here that:

* A K/D with approximately a K/D of 4.0 indicates a high degree of skill.
* Those with a K/D of 5.0 seem to be a stable pool of highly skilled players
* Those with a K/D of around 6.0+ seem be the primary outliers to focus on for potential cheating as well since they are well beyond many other players.
* Squad statistics might need a bit of a focus on the further end of the ratios at approximately 6.0+ since the max ratio hits 12.5, well beyond the other datasets.

```{r, echo=FALSE, warning=FALSE}
load(file="solo.RData")

#Plot zoomed in Squad statistics around the higher end of K/D
qplot(team$squad_KillDeathRatio, 
      xlab="Squad K/D ratio",
      xlim=c(4,12.5),
      geom="histogram",
      binwidth=.05, 
      fill=I("blue"), 
      col=I("red"))

```

The data, at least on an individual level is precise to a thousandths place value. as such some of the data seems to mesh together on a higher binwidth. However, with this concentration of values we can see that an extremely few amount of players achieve a squad K/D of 8.0 and beyond. A stable pool, like the solo and duo datasets, appear around the 4.0 to 5.0 K/D range. How the values at apprximately 6.0+ are achieved should be put under scrutiny, since the pool is small enough to observe. 

### Win ratio visualization

```{r, echo=FALSE}
load(file="solo.RData")

#Solo K/D histogram
qplot(solo$solo_WinRatio, 
      xlab="Solo Win Ratio",
      geom="histogram",
      binwidth=.1, 
      fill=I("blue"), 
      col=I("red"))

#Duo K/D histogram
qplot(duo$duo_WinRatio, 
      xlab="Duo Win Ratio",
      geom="histogram",
      binwidth=.1, 
      fill=I("blue"), 
      col=I("red"))

#Squad K/D histogram
qplot(team$squad_WinRatio, 
      xlab="Squad Win Ratio",
      geom="histogram",
      binwidth=.1, 
      fill=I("blue"), 
      col=I("red"))

```

Notes:

* The win ratio average goes up depending on team composition: Solo<Duo<Squad
    + This makes sense considering that a victory a team extends to all members, including those that did not make it to the end alive.
*

## Data Modeling (Machine Learning)

